PyPy's Approach to Virtual Machine Construction

### 2006-07-13 talk/dls2006/paper.tex
a few sentence refactorings and a couple of XXX comments

There are two major components in PyPy:
%
\begin{enumerate}
\item the \textit{Standard Interpreter}: an implementation of the Python programming
language, mostly complete and compliant with the current version of the
language, Python 2.4.
\item the \textit{Translation Process}: a translation tool-suite whose goal is to
compile subsets of Python to various environments.
\end{enumerate}
%
In particular, we have defined a subset of the Python language called
``restricted Python'' or RPython.  This sublanguage is not restricted
syntactically, but only in the way it manipulates objects of different
types.  The restrictions are a compromise between the expressivity and
the need to statically infer enough type information to generate
efficient code.  RPython still supports exceptions, inheritance but
limited to single inheritance with some mix-in support, dynamic
dispatch, to some extent keywords arguments and varargs, first-class
function and class values, limited use of bound methods, runtime
\texttt{isinstance} and type queries, but no runtime
reflection. Bindings in class and global namespaces are assumed
constant. RPython code can be run on a Python interpreter without
severe semantics mismatches, figure \ref{fig_create_frame} shows some
RPython code from the \textit{Standard Interpreter}, it is still quite
idiomatic Python code using the built-in dictionary type and
first-class class values instantiation.

@@

@@

@@

It is also simple, in the context of abstract interpretation, to
perform translation-time constant propagation, computations and
implement conditional code in a natural way. For example, the code in
figure \ref{fig_insertclean}, which is then specialized for the
possible types of dictionary \texttt{d}, shows this kind of
compile-time computation and introspection on low-level types,
\texttt{ENTRY} is a C-like structure type and the
\texttt{hasattr}-ibute checks introspect it at translation time. This
kind of expressiveness increased our ability to reuse code even in
low-level helpers.

@@

>>>>

There are two major components in PyPy:
%
\begin{enumerate}
\item the \textit{Standard Interpreter}: an implementation of the Python programming
language, mostly complete and compliant with the current version of the
language, Python 2.4.
\item the \textit{Translation Process}: a translation tool-suite whose goal is to
compile subsets of Python to various environments.
\end{enumerate}
%
In particular, we have defined a subset of the Python language called
``restricted Python'' or RPython.  This sublanguage is not restricted
syntactically, but only in the way it manipulates objects of different
types.  The restrictions are a compromise between the expressivity and
the need to statically infer enough type information to generate
efficient code.  RPython still supports exceptions, inheritance but
limited to single inheritance with some mix-in support, dynamic
dispatch, to some extent keywords arguments and varargs, first-class
function and class values, limited use of bound methods, runtime
\texttt{isinstance} and type queries, but no runtime
reflection. Bindings in class and global namespaces are assumed
constant. RPython code can be run on a Python interpreter without
severe semantics mismatches. Figure \ref{fig_create_frame} shows some
RPython code from the \textit{Standard Interpreter} and it can be seen
that it is still quite idiomatic Python code using the built-in
dictionary type and first-class class values instantiation.

@@

%%% XXX can we be more optimistic about the OOTyper?  It's pretty
%%% likely that pypy.net will work in some form by OOPSLA -- hi
%%% Antonio :)

@@

%%% XXX if we're looking to save space, i think this reason can go,
%%% the others seem more important (to me, at least)

@@

It is also simple, in the context of abstract interpretation, to
perform translation-time constant propagation, computations and
implement conditional code in a natural way. For example, the code in
figure \ref{fig_insertclean}, which is then specialized for the
possible types of dictionary \texttt{d}, shows this kind of
compile-time computation and introspection on low-level types --
\texttt{ENTRY} is a C-like structure type and the
\texttt{hasattr}-ibute checks introspect it at translation time. This
kind of expressiveness increased our ability to reuse code even in
low-level helpers.

@@

%%% XXX the first sentence seems to come a bit out of the blue.
